---
title: "Snow Data Assignment: Web Scraping, Functions, and Iteration"
author: "Kate Weimer"
date: "2-10-2022"
output: html_document
---

```{r setup3, include=FALSE}
library(rvest)
library(tidyverse)
library(lubridate)
library(readxl)
library(rvest)

```

# 3: Webscraping and Iterations

## Assignment:

1. Extract the meteorological data URLs. Here we want you to use the `rvest` package to get the URLs for the `SASP forcing` and `SBSP_forcing` meteorological datasets.c

```{r, message = FALSE}
datapath = 'data/'
dir.create(datapath)

```


```{r}
site_url <- 'https://snowstudies.org/archived-data/'

#Read the web url
webpage <- read_html(site_url)

links <- webpage %>%
  html_nodes('a') %>%
  .[grepl('forcing',.)] %>%
  html_attr('href')

#Grab only the name of the file by splitting out on forward slashes
splits <- str_split_fixed(links,'/',8)

#Keep only the 8th column
dataset <- splits[,8] 

# view(dataset)

file_names <- paste0(datapath,dataset)

```


2. Download the meteorological data. Use the `download_file` and `str_split_fixed` commands to download the data and save it in your data folder. You can use a for loop or a map function. 

```{r, message = FALSE}
#generate a file list for where the data goes
file_names <- paste0('data/',dataset)

for(i in 1:2){
  download.file(links[i],destfile=file_names[i])
}

downloaded <- file.exists(file_names)

evaluate <- !all(downloaded)
```

3. Write a custom function to read in the data and append a site column to the data. 

```{r}

# this code grabs the variable names from the metadata pdf file
library(pdftools)
headers <- pdf_text('https://snowstudies.org/wp-content/uploads/2022/02/Serially-Complete-Metadata-text08.pdf') %>%
  readr::read_lines(.) %>%
  trimws(.) %>%
  str_split_fixed(.,'\\.',2) %>%
  .[,2] %>%
  .[1:26] %>%
  str_trim(side = "left")

```


```{r}
read_in_weatherdata <- function(file){
name = str_split_fixed(file,'_',3)[,2] 
   df <- read.delim(file, header = F, sep = "", skip =4) %>%  mutate(site = name)
return(df)
}

```


4. Use the `map` function to read in both meteorological files. Display a summary of your tibble.
```{r, message= FALSE}
setwd("~/Spring 2022/ESS580/3_snow_functions_iteration/data")
weather_data_full <- map_dfr(dataset, read_in_weatherdata) 

weather_data_full <- select(weather_data_full, V1, V2, V10, site)%>% rename(Year = V1, Month = V2, temp = V10)

summary(weather_data_full)

```

5. Make a line plot of mean temp by year by site (using the `air temp [K]` variable). Is there anything suspicious in the plot? Adjust your filtering if needed.

```{r, warning= FALSE, message= FALSE}
weather_data_yearlymean <- weather_data_full %>% group_by(Year,site) %>% filter(Year != 2003, Year != 2011) %>% summarize(meantemp = mean(temp))
```


```{r}
ggplot(weather_data_yearlymean, aes(x = Year, y = meantemp, color = site )) +
  geom_line() +
  labs( x = "Year", y = "Mean Temperature (K)")
```

The dataset only included parts of 2003 and 2011, which skewed the yearly average for both years. 

6. Write a function that makes line plots of monthly average temperature at each site for a given year. Use a for loop to make these plots for 2005 to 2010. Are monthly average temperatures at the Senator Beck Study Plot ever warmer than the Snow Angel Study Plot?
Hint: https://ggplot2.tidyverse.org/reference/print.ggplot.html


```{r}
# create a function
yearly_plotter <- function(df, year){
  
  monthly_mean <- df  %>% 
    group_by(Year, Month, site) %>% 
    summarize(meantemp = mean(temp)) %>% filter(year == Year) 
  
   
  figure <- ggplot(monthly_mean, aes(x = Month, y = meantemp, color = site )) +
  geom_line() +
  labs( x = "Month", y = "Mean Temperature (K)") 

  
  print(figure)
}
```

```{r, message= FALSE}
# run the function in a for loop

x <- c(2005, 2006, 2007, 2008, 2009, 2010)

for(year in x){(yearly_plotter(weather_data_full, year))}
```

