---
title: "LAGOS Spatial Analysis"
author: "Kate Weimer"
date: "23 February 2022"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
```


# 4: LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)

```{r data-read, warning=FALSE}
# #Lagos download script
#LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path())
#Load in lagos
lagos <- lagosne_load()

#Grab the lake centroid info
lake_centers <- lagos$locus

```



### Convert to spatial data
```{r}
#Look at the column names
#names(lake_centers)

#Look at the structure
#str(lake_centers)

#View the full dataset
#View(lake_centers %>% slice(1:100))

spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

#Subset for plotting
subset_spatial <- spatial_lakes %>%
  slice(1:100) 

subset_baser <- spatial_lakes[1:100,]

#Dynamic mapviewer
mapview(subset_spatial)

```


### Subset to only Minnesota

```{r}
states <- us_states()

#Plot all the states to check if they loaded
#mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  arrange(-lake_area_ha) %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')
```



## 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r}
IA_IL <- states %>%
  filter(name == "Iowa" | name == "Illinois") %>%
  st_transform(2163)


mapview(IA_IL)
```



## 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa combined? How does this compare to Minnesota?

```{r}
IA_IL_lakes <- spatial_lakes[IA_IL,]
count(IA_IL_lakes)
```

There are 16466 sites in Iowa and Illinois combined, which is a little more that half the sites in Minnesota alone. 


## 3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on 
y axis (check out geom_histogram)


```{r}
Q3 <- spatial_lakes %>% filter(state_zoneid == "State_14" | state_zoneid == "State_13") %>% 
  mutate(state = ifelse(state_zoneid == "State_14", paste("Minnesota"), paste("Iowa")))

ggplot(Q3, aes(lake_area_ha))+
  geom_histogram(bins = 4) +
  # scale_x_continuous(breaks = seq(0, 130000, 4))
  facet_wrap(~state) 
  
 max(Q3$lake_area_ha)
```

The majority of lakes in both states are under 25,000 acres. Minnesota has many more lakes.  

## 4) Make an interactive plot of lakes in Iowa and Illinois and color them by lake area in hectares

```{r}
IA_IL_lakes %>%
  arrange(-lake_area_ha) %>%
  mapview(.,zcol = 'lake_area_ha')
```


## 5) What other data sources might we use to understand how reservoirs and natural lakes vary in size in these three states? 

Lake volume could be a more meaningful metric than lake area for people interested in available water for agriculture or other purposes. 
